\input{111pre}
\begin{document}
\MYTITLE{Examination 1 Study Guide \\ Delivered: Thursday, November 12, 2015 \\ Examination: Thursday, November 19, 2015, 11:00 am}

\section*{Introduction}

This course will have its first examination on Thursday, November 19, 2015 from 11:00 to 12:15 pm. The examination will
be ``closed notes'' and ``closed book'' and it will cover the following content. Please review the ``Course Schedule''
on the Web site for the course to see the content and slides that we have covered in the first module. You may post
questions about this material to Slack.

\begin{itemize}

  \itemsep 0in

  \item Chapters One through Three in SETP (i.e., introduction to the software engineering lifecycle)

  \item Chapters One through Three in MMM (i.e., challenges and solutions in software engineering)

  \item Chapters Four through Six in SETP (i.e., requirements, architecture, and design)

  \item Chapter Four in MMM (i.e., trade-offs in strategies for system design)

  \item Chapter Seven in SETP (i.e., software implementation)

  \item Chapters Twelve and Thirteen in MMM (i.e., tools for implementation and debugging)

  \item Your class notes, class activities, lecture slides, and the first four laboratory assignments

  \item Knowledge of the basic commands necessary for using {\tt git} and Bitbucket; basic understanding of the Markdown
    syntax and the use of associated command-line tools such as {\tt pandoc}

  \item Key concepts introduced and reviewed during the laboratory sessions (e.g., the components of a ``programming
    systems product'', strategies for performing next-release planning, and metrics for evaluating the quality of a
    system's design and implementation)

  \item Lessons learned from the team-based implementation and delivery of real-world software

\end{itemize}

\vspace*{-.05in}

\noindent The examination will include a mix of questions that will require you to draw and/or comment on a diagram,
write a short answer, explain and/or write a source code segment, or give and comment on a list of concepts or points.
The emphasis will be on the following list of illustrative topics; yet, this topic list is not exhaustive in nature.
While the examination is cumulative, the focus of this examination will be on the content that we have covered since the
last examination.

\vspace*{-.05in}
\begin{itemize}

  \itemsep 0in

  \item The state-of-the-art and the key challenges within the field of software engineering, with a focus on the steps
    of problem solving and the meaning of terms like ``defect'' and ``quality''.

  \item The phases of the software development lifecycle and the ways in which different software process models (e.g.,
    the spiral model or the V model) connect and interpret these phases.

  \item The key strengths and weaknesses of the different software development process models (e.g., one drawback of
    the waterfall model is its focus on documents and its lack of explicit iteration).

  \item Key terms such as ``verification'' and ``validation'' and ``incremental'' and ``iterative''.

  \item How to use activity graphs to track progress and plan a software development project. Additionally, an
    understanding of the ways in which managers will estimate the deadlines for completing a software system (e.g.,
    using data mining algorithms to predict project characteristics such as anticipated costs and the likelihood of an
    on-time completion).

  \item The roles that members of a software team may play and the ways in which individual personalities
    and characteristics may equip certain people to work on specific tasks.

  \item How different types of software engineering tasks exhibit different relationships between the time-to-completion
    and the number of workers assigned to finish the task.

  \item Lessons learned from working in a team to specify, design, implement, test, document, and release a programming
    systems product during our laboratory sessions. For instance, an understanding of the technical means by which a
    program's code and documentation is released to both Bitbucket and GitHub. In addition, a grasp of how to use
    GitHub's issue tracking system to report, triage, and resolve defects and/or concerns about a software system.

  \item The reasons why it is important to accurately capture the requirements for a software system.

  \item The different ways in which you can classify and understand software requirements.

  \item Characteristics of requirements and the ways in which they can be validated and verified.

  \item The relationship between the architecture and design of a software system.

  \item Canonical software architectures (e.g., ``pipe and filter'') and their strengths and weaknesses.

  \item Principles for improving and evaluating a system's design. For instance, in the context of coupling, an
    understanding of tight and loose coupling, the kinds of architectures that promote different types of coupling, and
    examples of metrics calculated by tools such a JDepend.

  \item The ways in which pre-conditions, post-conditions, and invariants can document an interface.

  \item A basic understanding of object-oriented design and the effective use of design patterns.

  \item The different types of Unified Modelling Language (UML) diagrams, the purpose, strengths, and weaknesses of each
    canonical type of diagram, and, overall, the purpose of using diagrams in software design. For instance, an
    understanding of the UML package and class diagrams.

  \item An understanding of the phrase ``conceptual integrity'' and, additionally, the ability to analyze different
    strategies for achieving conceptual integrity in all phases of the software lifecycle.

  \item The tension between ``aristocracy'' and ``democracy'' in the design of software systems.

  \item The meaning and purpose of various object-oriented design and implementation measures.

  \item Standards and guidelines that govern the way in which a program's source code should be implemented and
    formatted. A high-level understanding of a source code standard for the Java programming language and a clear sense
    of the benefits and drawbacks associated with using this standard. Additionally, the ability to look at a source
    code segment, written in Java, and comment on how it adheres to or violates a Java coding standard.

\end{itemize}

\vspace*{-.05in}
\noindent Minimal partial credit may be awarded for the questions that require a student to write a short answer. You
are strongly encouraged to write short, precise, and correct responses to all of the questions. When you are taking the
examination, you should do so as a ``point maximizer'' who first responds to the questions that you are most likely to
answer correctly for full points. Please keep the time limitation in mind as you are absolutely required to submit the
examination at the end of the class period unless you have written permission for extra time from a member of the
Learning Commons. Students who do not submit their examination on time will have their overall point total reduced.
Please see the course instructor if you have questions about any of these policies.

\vspace*{-.1in}
\section*{Review the Honor Code}
\vspace*{-.06in}

\noindent Students are required to fully adhere to the Honor Code during the completion of this examination. More
details about the Allegheny College Honor Code are provided on the syllabus. Students are strongly encouraged to
carefully review the full statement of the Honor Code before taking \mbox{this test}.

\noindent The following provides you with a review of Honor Code statement from the course syllabus:

The Academic Honor Program that governs the entire academic program at Allegheny College is described in the Allegheny
Academic Bulletin.  The Honor Program applies to all work that is submitted for academic credit or to meet non-credit
requirements for graduation at Allegheny College.  This includes all work assigned for this class (e.g., examinations,
laboratory assignments, and the final project).  All students who have enrolled in the College will work under the Honor
Program.  Each student who has matriculated at the College has acknowledged the following pledge:

\vspace*{-.11in}
\begin{quote}
  I hereby recognize and pledge to fulfill my responsibilities, as defined in the Honor Code, and to maintain the
  integrity of both myself and the College community as a whole.
\end{quote}
\vspace*{-.11in}

\vspace*{-.15in}
\section*{Strategies for Studying}
\vspace*{-.05in}

As you study for this examination, you are encouraged to form study groups with individuals who were previously a
member of one of your software development teams during a laboratory session. You can collaborate with these individuals
to ensure that you understand all of the key concepts mentioned on this study guide. Additionally, students are
encouraged to create a Slack channel that can host questions and answers that arise as you are studying for the test.
Even though the course instructor will try to, whenever possible, answer review questions that students post in this
channel, you are strongly encouraged to answer the questions posted by your colleagues as this will also help you to
ensure that you fully understand the material.

When studying for the test, don't forget that the Web site for our course contains mobile-ready slides that will provide
you with an overview of the key concepts that we discussed in the first module. You can use the color scheme in the
slides to notice points where we, for instance, completed an in-class activity, discussed a key point, or made reference
to additional details available in the SETP and MMM textbooks. Finally, students are strongly encouraged to schedule a
meeting during the course instructor's office hours so that we can resolve any of your questions about the material and
ensure that you have the knowledge and skills necessary for doing well on this examination. 

Remember, while the test is taken individually, your review for it can be done collaboratively!

\end{document}
